---
title: "R Analytics using ggplot2 "
output: html_notebook
---

The main difference is that, unlike base graphics, ggplot works with dataframes and not individual vectors. 

All the data needed to make the plot is typically be contained within the dataframe supplied to the ggplot().

```{r}
#install.packages(ggplot2)
library(ggplot2)
```

**mtcars dataset in R**

A data frame with 32 observations on 11 (numeric) variables.

[, 1]	mpg	Miles/(US) gallon

[, 2]	cyl	Number of cylinders

[, 3]	disp	Displacement (cu.in.)

[, 4]	hp	Gross horsepower

[, 5]	drat	Rear axle ratio

[, 6]	wt	Weight (1000 lbs)

[, 7]	qsec	1/4 mile time

[, 8]	vs	Engine (0 = V-shaped, 1 = straight)

[, 9]	am	Transmission (0 = automatic, 1 = manual)

[,10]	gear	Number of forward gears

[, 11] carb     number of carburetors

```{r}
data("mtcars")
head(mtcars)
```


**Scatter plot with ggplot**

```{r}
df = mtcars[ , c('cyl','mpg')]
ggplot(mtcars , aes(x=cyl, y=mpg)) +  geom_point(size = 4)
# wt vs hp , wt vs mpg, hp vs mpg
```


```{r}
df = mtcars[, c('mpg','hp')]
ggplot(df , aes(x=mpg, y=hp)) +  geom_point(size = 1)
```



**mpg dataset in R**


Variable	Type	Description	Details

manufacturer	string	car manufacturer	15 manufacturers

model	string	model name	38 models

displ	numeric	engine displacement in liters	1.6 - 7.0, median: 3.3

year	integer	year of manufacturing	1999, 2008

cyl		number of cylinders	4, 5, 6, 8

trans	string	type of transmission	automatic, manual (many sub types)

drv	string	drive type	f, r, 4, f=front wheel, r=rear wheel, 4=4 wheel

cty	integer	city mileage	miles per gallon

hwy	integer	highway mileage	miles per gallon

fl	string	fuel type	5 fuel types (diesel, petrol, electric, etc.)

class	string	vehicle class	7 types (compact, SUV, minivan etc.)



**Box plots**

 It visualises five summary statistics (the median, two hinges and two whiskers), and all "outlying" points individually.


```{r}
head(mpg)
ggplot(mpg, aes(class, hwy)) + geom_boxplot()
ggplot(mpg, aes(model, hwy)) + geom_boxplot()
ggplot(mpg, aes(class, hwy)) + geom_boxplot(outlier.colour = "red", outlier.shape = 2)
ggplot(mpg, aes(hwy, class)) + geom_boxplot()

```
 
```{r}
mpg$year
```
 
 

```{r}
ggplot(mpg, aes(factor(year), hwy)) +
 geom_boxplot(outlier.colour = "red", outlier.shape = 5)
```

```{r}
ggplot(mpg, aes("", hwy)) +
 geom_boxplot(outlier.colour = "red", outlier.shape = 5)
```



**Histogram **

```{r}
ggplot(data=mpg, aes(cty)) + geom_histogram(binwidth=1)
```


```{r}
ggplot(data=mpg, aes(displ)) + geom_histogram(binwidth=2)
ggsave("displacement.png")
```

```{r}
data(mtcars)
ggplot(mtcars,aes(cyl,mpg))+ geom_boxplot()
ggplot(mtcars,aes(mpg,cyl))+ geom_boxplot()
ggplot(mtcars,aes(factor(cyl),mpg))+ geom_boxplot()

```


```{r}
head(iris)
iris$Sepal.Width = iris$Sepal.Width *10 
iris$Petal.Length = iris$Petal.Length * 10
head(iris)
ggplot(iris,aes(x=Sepal.Width,y=Petal.Length))+geom_point(size=4) # error with quote
# if quotes are given
# data is taken as a single string
ggplot(iris,aes(x='Sepal.Width',y='Petal.Length'))+geom_point(size=4)

```

```{r}
ggplot(mpg,aes(x=drv,y=trans))+geom_point()

```


**Load Heart disease data set**

```{r}
mydata = read.csv("76_attributes_heartdiseases.csv",na.strings=c("-9","-18"))
names(mydata)
```

```{r}
min(mydata$V3)
max(mydata$V3)
mydata$age_group = cut( mydata$V3, breaks=c(29,50,77),labels=c("29-50","50-77"))
#mydata$age_group
ggplot(mydata,aes(x=age_group,y=V10))+geom_boxplot()
ggplot(mydata,aes(x="",y=V10))+geom_boxplot()
ggplot(mydata,aes(x="",y=V1))+geom_boxplot()
```

