---
title: "R Dataframes"
output: html_notebook
---

**Create Dataframe from matrix / array**

_as.data.frame()_

```{r}
M1 = matrix(c(1,2,3,4),nrow=2)
#M1
df <- as.data.frame(M1)
df
df$V1
```

```{r}
str(df)
```


```{r}
is.data.frame(df)
```

```{r}
nrow(df)
```

```{r}
ncol(df)
```


```{r}
length(df)
```

```{r}
colnames(df)
```

```{r}
names(df)
```


```{r}
M1 = matrix(c(1,2,3,4),nrow=2)
m_df <- as.data.frame(M1)
colnames(m_df) = c('c1','c2') # change column names
print(m_df)
str(m_df)
ncol(m_df)
```


```{r}
#nominal varible doesn't have any order
emp = c("John Doe", "Peter Gynn", "Jolie Hope")
#Ordinal varible has order
designation <- factor(c("Manager", "Team Lead","SME"), ordered =TRUE,          levels = c("SME", "Team Lead","Manager"))
salary <- c(41000, 33400, 26800)
startdate <- as.Date(c("2010-11-1", "2008-3-25", "2007-3-14"))
employee.data <- data.frame(emp, designation,salary, startdate)

str(employee.data)
employee.data
```

```{r}
?data.frame
```


```{r}
employ.data <- data.frame(emp, salary, startdate, stringsAsFactors=TRUE)
str(employ.data)
```

```{r}
# variaable name can have .
employ.data <- data.frame(emp, salary, startdate, stringsAsFactors=FALSE)
# select rows based on condition
employ.data[employ.data$salary>35000,]
```

```{r}
employ.data$salary
```

```{r}
employ.data[c('emp','salary')]
```

```{r}
employ.data[employ.data$salary>35000,c('emp','salary')]
```


**Reading TEXT / CSV into Data frame**

read.csv(file, header = TRUE, sep = ",", quote = "\"",
         dec = ".", fill = TRUE, comment.char = "", ...)

GET HELP --> ?read.csv

read.csv2(file, header = TRUE, sep = ";", quote = "\"",
          dec = ",", fill = TRUE, comment.char = "", ...)

```{r}
df = read.csv("76_attributes_heartdiseases.csv")
summary(df)
```


```{r}

#df_sample = read.csv("sample3.txt",sep="")
#df_sample
df_sample = read.csv("sample1.txt",sep="|")
df_sample
```

```{r}
df = read.csv("76_attributes_heartdiseases.csv")
summary(df$V73)
df = read.csv("76_attributes_heartdiseases.csv",na.strings=c("-9","-18"))
summary(df$V73)
```

```{r}
summary(df)
```


```{r}
df_sample = read.csv("sample2.txt",skip=2,sep="|")
df_sample
df_sample = read.csv("sample2.txt",skip=2,sep="|",blank.lines.skip=FALSE)
df_sample
df_sample = read.csv("sample2.txt",skip=0)
df_sample
```

**Few more IMP operations in read.csv**

stringsAsFactors --> Default to TRUE. String columns will be treated as factors

                    If FALSE then Strings will be kept intact

encoding --> For handing unicode or other encoding problems

nrows --> integer: the maximum number of rows to read in. 

            Negative and other invalid values are ignored.
            
            Useful when loading very large files 



```{r}
print(ncol(df))
```

```{r}
print(nrow(df))
```

```{r}
summary(df)
```

**Create subset of a dataframe**

We can use subset() function. HELP --> ?subset

Arguments:

data / data frame

condition based on which to create the subset

select --> specify the array of columns to be selected


```{r}
df_1 = subset(df, df$V1 > 200)
head(df_1)
df_1 = subset(df, df$V1 > 200, select = c("V1","V2"))
head(df_1)
nrow(df_1)
```


```{r}
head(mtcars)
mtcars[mtcars$mpg > 20 , ]
mtcars[mtcars$mpg > 20 , c('mpg')]
mtcars[(mtcars$mpg > 20) & (mtcars$cyl != 6), c('wt','cyl','mpg')]


```

**Select all Numeric Columns or All String/Character Columns**

```{r}
library("ggplot2")
head(mpg)
Filter(is.numeric, mpg)
Filter(is.character, mpg)
```

```{r}
M1 = matrix(c(1234,2235,67,85),nrow=2)
m1_df <- as.data.frame(M1)
colnames(m1_df) = c('roll_no','marks_mod1')
m1_df
M2 = matrix(c(1234,2235,75,68),nrow=2)
m2_df <- as.data.frame(M2)
colnames(m2_df) = c('roll_no','marks_mod2')
m2_df
merge(m1_df,m2_df,by = "roll_no")
```

**Merge Dataframes**
```{r}
M1 = matrix(c(1234,2235,67,85),nrow=2)
m1_df <- as.data.frame(M1)
colnames(m1_df) = c('roll_no','marks_mod1')
m1_df
M2 = matrix(c(1234,2235,75,68),nrow=2)
m2_df <- as.data.frame(M2)
colnames(m2_df) = c('r_no','marks_mod2')
m2_df
M3 = matrix(c(1234,2235,50,45),nrow=2)
m3_df <- as.data.frame(M3)
colnames(m3_df) = c('r_no','marks_mod3')
m3_df
merge(m1_df,m2_df,by.x = "roll_no", by.y = "r_no")
merge(m2_df,m3_df,by = "r_no")
```

**Store Datframe to a File **

```{r}
write.csv(m2_df,"new_df.csv")
```

