---
title: "R dplyr"
output: html_notebook
---

```{r}
#install.packages("dplyr")
library(dplyr)
```

```{r}
# Clear R workspace
rm(list = ls() ) 
```

```{r}
movies <- read.csv("C:/Users/dbda/Documents/moviesData.csv")
movies
```

**dplyr filter function**

```{r}
moviesComedy <- filter(movies,
                       genre == "Comedy")
moviesComedy
```

```{r}
moviesComDrP <- filter(movies,
                       genre %in% c("Comedy", "Drama"))
moviesComDrP
```


```{r}
moviesComIm <- filter(movies,
                       genre == "Comedy" & imdb_rating >= 7.5)
moviesComIm
```

```{r}
moviesRunRel <- filter(movies,runtime < 100 | thtr_rel_year > 2000)
moviesRunRel
```


**dplyr arrange**

```{r}
# Arrange the movies in ascending order of imdb_rating
moviesImA <- arrange(movies, imdb_rating) 
moviesImA
```





```{r}
# Arrange the movies both by genre and imdb_rating
moviesGeIm <- arrange(movies, genre, imdb_rating)
moviesGeIm
```

```{r}
levels(movies$mpaa_rating)
```

```{r}
# Arrange the movies in descending order 
moviesImD <- arrange(movies, runtime, desc(mpaa_rating),genre) 
moviesImD
```

**dplyr select operation**

```{r}
moviesTGI = select(movies, title, genre, imdb_rating)
moviesTGI
```

```{r}
moviesTtoI = select(movies, title:imdb_rating)
moviesTtoI
```

```{r}
moviesTtoI = select(movies, c(2,5,7)) # -c(4,10) c(2,5,7) 2,5,7
moviesTtoI
```

**select function options**

starts_with
ends_with --> select all columns ends with 'rating'
contains 
matches --> find macthing columns using regular expressions "?*.[]"
num_range("y",1:5) --> y1,y2,y3,y4,y5  
one_of --> select one of the columns of array
any_of

```{r}
names(movies)
```



```{r}
moviesTHT = select(movies, title, starts_with("best"))
names(moviesTHT)
```

```{r}
moviesTHT = select(movies, title, contains("rel"))
names(moviesTHT)
```


```{r}
moviesTHT = select(movies, title, contains("imdb"))
names(moviesTHT)
# match with Imdb imdb
moviesTHT = select(movies, title, matches("[Ii]mdb")) 
names(moviesTHT)

```

```{r}
movies_acto_15 = select(movies, num_range("actor",1:5))
names(movies_acto_15)
```



```{r}
hd_df = read.csv("76_attributes_heartdiseases.csv")
hd_df_v1_v5 = select(hd_df, num_range("V",10:5))
names(hd_df_v1_v5)
```

```{r}
# any_of is used to remove the columns efficiently
moviesTHT = select(movies, - any_of(c("title","imdb_rating")))
names(moviesTHT)
moviesTHT = select(moviesTHT, - any_of(c("title","imdb_rating")))
names(moviesTHT) # no error even if columns are absent

```

**dplyr rename**

```{r}
moviesR = rename(movies, rel_yr = "thtr_rel_year") # new_name = "existing name"
moviesR
```

**dplyr mutate**

```{r}
moviesMu = mutate(movies, 
                  newScore = max(critics_score - audience_score,0),
                  avg_Score = (runtime + critics_score + audience_score) / 3,
                  min_score = min(critics_score,audience_score)
                  )
moviesMu
moviesMu$avg_Score
```

```{r}
# Print all movies which are released after 2001 , genre drama and avg_score > 60
# where avg_score = (critics_score+audience_score)/2

moviesnew=filter(movies,thtr_rel_year>2001,genre=="Drama")
moviesnew
moviesmutatenew=mutate(moviesnew,avg_score=(critics_score+audience_score)/2)
moviesfinal=filter(moviesmutatenew,avg_score>60)
moviesfinal
```

**Exercise**

Print all movies which have runtime > 100, in descending order of new_rating 

new_rating = (critics_score+audience_score)/ imdb_num_votes





```{r}
mv = filter(movies,actor1 == "John Cusack" )
arrange(mv,desc(imdb_rating))
```


**dplyr summarise**

min(x) - minimum value of vector x .
max(x) - maximum value of vector x .
mean(x) - mean value of vector x .
median(x) - median value of vector x .
quantile(x, p) - pth quantile of vector x .
sd(x) - standard deviation of vector x .
var(x) - variance of vector x .
IQR(x) - Inter Quartile Range (IQR) of vector x .
diff(range(x)) - total range of vector x .

special functions

first(x) - The first element of vector x .
last(x) - The last element of vector x .
nth(x, n) - The nth element of vector x .
n() - The number of rows in the data.frame or group of observations that summarise()
describes.
n_distinct(x) - The number of unique values in vector 


```{r}
moviesComDrP <- filter(movies,
                       genre %in% c("Comedy", "Drama"))
summarise(moviesComDrP, mean(imdb_rating))
summarise(movies,mean_imdb =  mean(imdb_rating), qt = quantile(imdb_rating,0.5))
```

```{r}
mv = filter(movies,actor1 == "John Cusack" )
summarise(mv, max(imdb_rating))
```
```{r}
summary(movies$imdb_num_votes) #na.rm=TRUE
```
```{r}
summarise(movies, min(imdb_num_votes),max(imdb_num_votes),
          mean(imdb_num_votes),median(imdb_num_votes),
          sd(imdb_num_votes),quantile(imdb_num_votes,0.2)
          ,n_distinct(imdb_num_votes),n())

```


**dplyr group_by**

```{r}
grpMovies = group_by(movies,genre)
grpMovies
```

```{r}
moviesDr <- filter(movies,
                       genre == "Drama")
grpMoviesDr = group_by(moviesDr,mpaa_rating)
summarize(grpMoviesDr,mean(imdb_rating))
```

```{r}
groupmovies=group_by(movies,genre)

summarise(groupmovies,mean(audience_score),mean(imdb_rating),median(imdb_num_votes))

```

**dplyr pipe operator**

%>% pipe operator

```{r}
movies %>%
  filter(genre == "Drama") %>%
  group_by(actor1) %>%
  summarize(mean_act_rating = mean(imdb_rating)) %>%
  arrange(desc(mean_act_rating)) %>%
  head()
# get best 6 actors which have high mean imdb ratings
```



```{r}
# WAP to get first 10 actors in drama movies with highest mean imdb rating
movies %>%
  filter(genre == "Drama") %>%
  group_by(actor1)         %>%
  summarize(mean_act_rating = mean(imdb_rating)) %>%
  ungroup() %>% # release the group
  arrange(desc(mean_act_rating)) %>%
  slice(1:10)#slice((n()-9):n()) #slice_tail(n=10)

```

**Store output in the Dataframe**

```{r}
actor_rating =   movies %>%
  filter(genre == "Drama") %>%
  group_by(actor1)         %>%
  summarize(mean_act_rating = mean(imdb_rating)) %>%
  ungroup() %>%
  arrange(desc(mean_act_rating)) %>%
  as.data.frame()

actor_rating
```

```{r}
df = read.csv("cleaned_76hd.csv")
head(df)
```

```{r}
static_m_f=df %>%
  group_by(V4,V58)%>%
  summarise(median(V9),mean(V12),mean(V10),mean(V40))

static_m_f
```
```{r}
static_s_nos=df %>%
  group_by(V13,V58)%>%
  summarise(n(),median(V9),mean(V12),mean(V10),mean(V40))

static_s_nos
```

```{r}

# WAP  with PIPE on employeeattrition.csv
#1. count employees with highest hourly rate
#2. count employees who stay near the company 
# and who left (attrition =1) the company  
#3. Find avg hourly rate of every department
#4. Find employees in any one department with daily rate more than average

```


```{r}
df = read.csv("EmployeeAttrition.csv")
ncol(df)
names(df)
```

```{r}
max(df$HourlyRate)
result=summarise(df,max(HourlyRate))
result

```

```{r}
df %>%
  filter(HourlyRate == max(HourlyRate)) %>%
  summarise(n())

```

```{r}
#3. Find avg hourly rate of every department
df %>%
  group_by(Department)%>%
  summarise(mean(HourlyRate))

```


```{r}
#4. Find employees in any one department with daily rate more than average
# Department == 'Sales'
df %>%
  filter(DailyRate > mean(DailyRate), Department == 'Sales') 
```

