---
title: "Basic R"
output: html_notebook
---

**Print Hello world**

These are different ways to print

```{r}
print("Hello world !!")
cat("... This is "," another ", "way!")
cat("\n But '\\n' is must!!")
paste("still another way ...", "..great!","kll") # 
print(paste("still another way ...", "..great!"))
```

**Accept user input**


```{r}
yourname <- readline(prompt="What is your name? ")
print(yourname)
print(class(yourname))
age <- readline(prompt="What is age? ")
print(age)
print(class(age))
```

**Setting current Working Directory**


```{r}
# \\ /
setwd("D:/R_projects/CDAC_Statistical_analysis_with_R")
```




**All variables of Environment**
When you want to remind yourself of all the variables you’ve created in
the environment, use ls().

```{r}
ls()
```

**Deleting Variables**
Variable can be deleted using rm()
You can use any variable name you have created

```{r}
ls()
rm(yourname)
ls()
```


**Creating variables and Array**

Variable types are 
(As per the oder)
Logical
Integer
Numeric
Complex
Character
** list

```{r}

x=100000L
print(class(x))
x=10
print(class(x))
x=10.67
print(class(x))
x=TRUE
print(class(x))
x=5+4i
print(class(x))
x="Hi"
print(class(x))

```

**Create Array**

In R we can use c() function also known as combine.

```{r}
x_dim1 = c(10,20,30,40,50,60)
print(x_dim1)
```



```{r}
x_dim1 = c(10,20,30,40,50,60)
print(x_dim1)
x_mix = c(3,0,TRUE,2+2i) #3+0i 0+0i 1+0i 2+2i
print(x_mix)
x_mix = c(0,"IACSD R Prog",TRUE,2+2i) #"0"            "IACSD R Prog" "TRUE"         "2+2i"   
print(x_mix)



############# VECTOR Function
x_vec = vector("numeric", 5) 
print(x_vec)
x_vec = vector("complex", 5)
print(x_vec)
x_vec = vector("logical", 5)
print(x_vec)
x_vec = vector("character", 5)# array of strings
print(x_vec)
x_vec = vector("list", 5)
print(x_vec)

```

```{r}
############# ARRAY Function
x_dim2 = array(c(10,20,30,40,50,60,70),dim=c(2,3))
print(x_dim2)
```

```{r}
# WAP to create array for 2 by 2
```



```{r}
# WAP to create array for 3 by 4
x_dim2 = array(c(1,2,3),dim=c(3,4))
print(x_dim2)
```

```{r}
?array
```



```{r}
x_dim2 = array(c(10,20,30,40,50),dim=c(2,2))
print(x_dim2)
```



**Creating Sequences by ':' operator**

```{r}
print(3:12)
print(8.5:4.5)
print(-12:1)
print(c(1, 1:3, c(5, 8), 13))
```

```{r}
print(10.45:5)
a = c(10:20)
a

```

```{r}
print(c(1, 1:3, c(5, 8), 13))
```



```{r}
x_dim2 = array(1:9,dim=c(3,3))
print(x_dim2)
```


**Creating Sequences by seq class**

Similar to range function in python

```{r}
print(seq.int(3, 12))
print(seq.int(3, 12, 0.5))
print(seq.int(0.1, 0.01, -0.01))
print(seq.int(2,100,2))
print(seq.int(99,90,-2))
```

```{r}
print(seq.int(10,20,-1))
```

**Sequence by seq_len function**

It takes 'n' as parameter. And creates sequence from 1:n

```{r}
print(seq_len(10))
```

**Sequence by seq_along function**

It is useful when we want to create sequence of length of input array

```{r}
v = c(45,56,66,34,23,100,450)
print(v)
seq_along(v)
print(seq_len(length(v)))
```

**Creating Vectors by Repeatition using rep() function**

```{r}
rep(1:5, 3)
rep(1:5, each = 3)
rep(1:5, times = 1:5)
rep(1:5, length.out = 7)
rep.int(1:5, 3)
rep_len(1:5, 7)
```

```{r}
rep(1:5, times = c(5,3,8,1,1))
# first 5 times ,, 2nd 3 times , 3rd 8 times, 4th 1 , 5 th 1
```


**Sum function**
```{r}
x=c(1,2,3)
sum(x)
x_dim2 = array(1:9,dim=c(3,3))
sum(x_dim2)
```



```{r}
r = array(c(10,67,-30,0,50,60),dim=c(2,3))
r
sum(r)
length(r)
max(r)
min(r)
```


**Length function**
```{r}
x_dim1
length(x_dim1)
x_dim2
length(x_dim2)
```

**Modify the length of Array**

```{r}
x = c(10,20,30,40,50)
#x = c(34:89)
print(x)
length(x) = 3
print(x)
x = c(10,20,30,40,50)
length(x) = 10
print(x)
x_vec = vector("numeric", 5)
length(x_vec) = 10
print(x_vec)
```


**Modify the elements in an Array**

```{r}
x_vec = vector("numeric", 5)
x_vec
x_vec[3] = 100
x_vec

x_vec = vector("numeric", 5)
x_vec
x_vec[1:3] = 89
x_vec

x_vec = vector("numeric", 5)
x_vec
x_vec[c(1,3,5)] = 99
x_vec

```



**Apply function**
Apply function has three arguments:: X, MARGIN and FUN.

X: is input array / data
MARGIN: MARGIN=1, it applies over rows, whereas with MARGIN=2, it works over columns.
FUN: Function to be applied on complete array / data of values. This function can be system defined or user defined.

```{r}
# Use apply to calculate the sum of the rows across all the matrices.
x_dim2 = array(1:12,dim=c(4,3))
print(x_dim2)
result <- apply(x_dim2, c(1), length)
print(result)
result <- apply(x_dim2, c(1), sum)
print(result)
result <- apply(x_dim2, c(2), sum)
print(result)

```

** Apply function on 3D array **

```{r}
# Use apply to calculate the sum of the rows across all the matrices.
x_dim3 = array(1:12,dim=c(2,2,2))
print(x_dim3)
result <- apply(x_dim3, c(1), length)
print(result)
result <- apply(x_dim3, c(1), sum)
print(result)
result <- apply(x_dim3, c(1,2), sum)
print(result)
```



```{r}
# WAP to create 3 by 3 array
# Print sum of each row and each column
x=array(seq.int(10,30,3),dim=c(3,3))
print(x)
result = apply(x, 1, sum)
print(result)
result = apply(x, 2, sum)
print(result)
c(apply(x, 1, sum),apply(x, 2, sum))
```

```{r}
# Create 4,3 array
# find maximum element in each row  -> max()
# find min element in each column-> min()

```


```{r}
x1 = array(c(1:9))
x1
sum(x1)
apply(x1, 1, sum)
x2 = array(c(1:9), dim=c(9,1))
x2
apply(x2, 1, sum)
apply(x2, 2, sum)
sum(x2)
```



**Names to array Elements **

```{r}
x = c(apple = 1, "banana" = 2, "kiwi fruit" = 3, 4)
print(x)
x <- 11:14  # ->
names(x) <- c("First", "Second", "Third", "Fourth")
print(x)
print(names(x))
```

**Indexing in Vectors**

```{r}
# vectorised operation , operation performed on every element of the array
x <- (1:10) ^ 2 

print(x)
x[2]
x[c(1,5,7)]
x[c(8,5,7)]
x[1:3]
x[-2] # exclude element at index 2
x[c(-1,-4)] 
#x[c(-1,-4,5)] #--> gives error
x[c(TRUE, TRUE, FALSE)] # repeated the same array
x[100]
x[-4.99] # floor rounding off
```

**Indexing in 2D array**

```{r}
x=array(seq.int(10,30,3),dim=c(3,3))
print(x)
x[5]
x[2,1] # NOT using c() --> it is index in 2D array
x[c(3,1),c(2,1)] # using C() --> it is index in 1D array
x[c(3,1),]

```


```{r}
x <- (1:5) ^ 2
x
x[c(TRUE, FALSE, FALSE, TRUE)]
x[c(TRUE)]
x[c(TRUE, FALSE)] # Broadcasting
```

**How to use condition to select elements from an Array**

```{r}
x <- (1:5) ^ 2
x
x[c(TRUE, FALSE, TRUE, FALSE, TRUE)]
x>10 # vectorized operations
x[x>10]
x[!x>10]
```

**Which Function**

```{r}
x <- (1:5) ^ 2
x
print(which(x>5)) # return index of elements which satisfy the condition
print(x[which(x>5)])
print(which.min(x)) # index of minimum element in array
print(x[which.min(x)])
print(which.max(x)) # index of max element
print(x[which.max(x)])
```


**Vector Operation**

Adding scalar to vector will add to all elements of that vector.

```{r}
x = 1:5
print( x + 1)
```

Adding two vectors of same length will do element wise addition

```{r}
x = 1:5
x1 = 11:15
print(x+x1)
```

Adding two vectors of different length will repeat small vector over longer vector

```{r}
x = 1:5
y = 1:10
print(x)
print(y)
print(x+y)
print(x-y)
print(x*y)
```

```{r}
x = 1:5
y = 11:18
x
y
x+y
```


**Deleting Elements from an Array**

There is no direct function to delete elements.

We can exclude the elements by using negative index as we did earlier.

```{r}
x = c(10,20,30,40,50)
print(x)
print("Delete second element")
x = x[-2]
print(x)
print("Delete multiple elements, index 1 and 4 of new array")
x = x[c(-1,-4)]
print(x)
```

**Cleaning the Environment**

```{r}
ls()
rm(list = ls())
ls()
```


**Controlling Visibility of Variables**

This is way to use what variables are shown in current environment.

```{r}
rm(list = ls())
x=10

print("List after creating x")
ls()

.xyz= 90

print("List after creating .xyz")
ls()

print("List of all variables")
ls(all.names = TRUE)
```
