---
title: "R Functions"
output: html_notebook
---

**User Defined Function**

These are used for customization by the user.
Repeated set of instructions are put in function to be used again
It makes programs readable and easy to understand

**Without argument**
Here function is created without any arguments.
It will have process inside it.
This function may or may not return any value

```{r}
display = function(){
  print("This is a function ....")
}
display
display() # calling the function
```



**With argument**

```{r}
add = function(a,b){ 
  # all functions can handle array / matrix/a single value/any data type
  return(a+b)
}
add(23,45)
add(c(3,4),c(6,7))
add(3+8i,7+4i)
add("k","l")

```

```{r}
# Create function to calculate tax for given income
# if income > 5 lac then tax = 20% of income
# else tax 10% of income
# return amount of tax
tax_calc = function(income){
  if(income>500000){
    tax=income*20/100
    #print(tax)
  }else{
    tax=income*10/100
    #print(tax)
  }
  return(tax)
}
tax_calc(900000)
apply(array(c(200000,80000,560000,45000),dim=c(4,1)),c(1) , tax_calc)
```


**With Default argument**

```{r}
pow = function(n,p=2){
  return(n^p)
}
pow(23)
pow(34,5)
```

**Lazy Evaluation of Function**

```{r}
square = function(a,b){
  print(a^2)
}
square(2)
square(2,56)    
```

```{r}
# WAP to create a function to 
# calculate total salary of an employee
# given basic
# HRA = 10 % of basic
# TA = 20% of basic
# total = basic + HRA + TA
# return total salary

```



**Built in functions**

```{r}
x = c(45,-34,23,11.33)
abs(x)
x = c(1,4,6)
log(x)
x = c(1,4,6)
exp(x) # base is e .. e^no
x = c(1,4,6)
sqrt(x)
x = c(1,4,6)
factorial(x)
x = c(1,4,6)
sin(x)
x = c(1,4,6)
cos(x)
x = c(1,0,-5,10,-34,0,0,100,-90)
sign(x)
```


```{r}
x=c(10,-20,-30,80,-90,40,20.1,0)
sign(x)
```

**Match Function**
match returns a vector of the positions of (first) matches of its first argument in its second.

%in% internally uses match

```{r}
x=c(10,20,30,10,20)
y=c(20,20,40)
match(x,y)
match(y,x)
ux <- unique(x)
ux
match(x, ux)
```

```{r}
x=c(10:30)
y=c(10,20,60)
x
y
match(y,x)

```

```{r}
x=c(10,20,30,40,50)
y=c(20,4,60,50,30)
x
y
match(x,y)
```



**Tabulate Function**

tabulate() function takes the integer-valued vector bin and counts the number of times each integer occurs in it

```{r}
tabulate(c(-5,0,3,5,4,5,4,10))
tabulate(c(3,5,4,5,4,100),nbins=4)
tabulate(c(3,5,4,8),nbins=4)
tabulate(c(3,5,4,8),nbins=10)
```


```{r}
x=c(10,300,30,10,300)
ux <- unique(x)
ux
match(x, ux)
tabulate(match(x, ux))
```



**NOTE ABOUT OPERATORS**

Every operator is a function

```{r}
'+'(90,45)
'*'(34,56)
```

