---
title: "R Loops"
output: html_notebook
---

**Repeat Loop**

The Repeat loop executes the same code again and again until specifically taken out by the programmer.


_NOTE_ If no specific _break_ statement is written, then this will be infinite loop.

```{r}
v = c("Inside", "loop")
cnt = 1
repeat{
  print(v)
  cnt = cnt+1
  if(cnt > 5){
    break
  }
}
```

**While Loop**

Loop runs till the test condition is TRUE. When test condition is FALSE then loop stops.

```{r}
cnt = 1
while (cnt <= 4)
{
  print(cnt)
  cnt = cnt + 1
}
```

**For Loop**


```{r}
print(LETTERS)
print(letters)
v <- LETTERS[21:length(LETTERS)]
for ( i in v) {
  print(i)
}
v1 = c(10,20,30)
for( i in v1){
  print(log(i))
}
log(v1)

```

Assignment::
Take an array
Print sum of alternate elements starting from index 2
ex a = c(90,67,78,34) ## 101

```{r}
a = c(90,67,78,34,23)
index=2
sum_a_alt=0
while( index <= length(a))
{
  sum_a_alt = sum_a_alt + a[index]
  index = index+2
}
print(sum_a_alt)
```

```{r}
a = c(90,67,78,34,23)
print(seq(2,length(a),2))
sum_a_alt=sum(a[seq(2,length(a),2)])
sum_a_alt
```


**Break statement**

_break_ statement will break the flow of the loop which is immediate. 

_NOTE_ If there is loop within loop then control will be taken out of only one loop.

```{r}
for ( i in 1:4)
  for (j in 10:25){
    print(j)
    if(j == 12)
      break
  }

```

**Next Statement**

The next statement in R programming language is useful when we want to skip the current iteration of a loop without terminating it.

```{r}
v = LETTERS[1:6]
for ( i in v){
  if (i == "D"){
    next
  }
  print(i)
}
```

**Making Loops Faster**

**lapply function**

```{r}
lapply
```

Example::
```{r}
# rnorm function generates random n values with mean 0 and std dev 1
x <- list(a = 1:5, b = rnorm(10))
x$b
lapply(x, mean)
```

**lapply with multiple arguments**



```{r}
x <- 1:4
lapply(x, runif, min = 10, max = 80)
```


**sapply (simplify lapply)**


sapply() calls lapply() on its input and then applies the following algorithm:


If the result is a list where every element is length 1, then a vector is returned


If the result is a list where every element is a vector of the same length (> 1), a matrix is returned.


If it can’t figure things out, a list is returned

```{r}
sapply
```


Example

```{r}
# rnorm function generates random n values with mean 0 and std dev 1
x <- list(a = 1:5, b = rnorm(10))
sapply(x, mean)
```

