---
title: "R Matrix Operations"
output: html_notebook
---

**Matrix**

Matrix can be created by specifying number of rows and columns.
"byrow = TRUE" means elements will be filled in row wise manner.

```{r}
M = matrix( c('a','a','b','c','b','a'), nrow=2,ncol=3,byrow = TRUE)
print(M)
```

If "byrow = TRUE" is not gievn then elements will be filled in column wise manner.
In this example elements 3 to 14 are arranged in 4 rows.  

```{r}
M <- matrix(c(1:12), nrow = 4)
print(M)
```

```{r}
M <- matrix(c(1:14), nrow = 4)
print(M)
```

In case rows or columns are not specified then, matrix with single column and no of rows equal to number of elements is created.

```{r}
M <- matrix(c(3:14))
print(M)
```



**Accessing elements of a Matrix**
```{r}
M <- matrix(c(3:14), nrow = 4)
print(M)
cat("\n Print element row 1 col 3 M[1,3]",M[1,3])
cat("\n Print row 1 M[1,]",M[1,])
cat("\n Print col 3 M[,3]",M[,3])
M[2,3] = 20  #Assigning value 20 to the element at 2nd row and 3rd column
cat("\n After assigning 20 to M[2,3] \n")
print(M)
M[3,] = 300 # This is vectorized ... all elements in row 3 will be 300
cat("\n After assigning 300 to row 3 \n")
print(M)

M[,2] = 23.4
cat("\n After assigning 23.4 to column 2 \n")
print(M)
```

```{r}
M <- matrix(c(3:14), nrow = 4)
print(M - 30) # create a copy of matrix
print(M) # original matrix is still the same
```



```{r}
M <- matrix(c(3:14), nrow = 4)
print(M)
M[2,2] = M[2,2] + 10
print(M)
M[3,] = M[3,] - 30
print(M)
M[,1] = M[,1] * 0.1
print(M)
M = M + 10101
print(M)

```


**Change Matrix Elements based on Condition**

```{r}
M <- matrix(c(3:14), nrow = 4)
print(M)
print(M[M < 6])
M[M < 6] = 100
print("Modified Array is ")
print(M)
M[M < 100] = M[M < 100] * 20
print("Modified Array is ")
print(M)
```

```{r}
# modify elements in 2nd row to 200 if element is less than 10
M <- matrix(c(3:14), nrow = 4)
M[2,][M[2,] < 10] = 200
M
```


**Add new column to matrix**

here we use 'cbind()' function

```{r}
M <- matrix(c(3:14), nrow = 4)
print(M)

M = cbind(M, c(0,0,0,0))

print("Modified matrix")
print(M)

M = cbind(M, c(50,51,52))
print(M)

M = cbind(M, -1)
print(M)
```

**Add new row to matrix**

```{r}
M <- matrix(c(3:14), nrow = 4)
print(M)

M = rbind(M , c(0))
print(M)

M = rbind(M , c(100,200,300,400))
print(M)

M = rbind(M , c(-1,-2))
print(M)
```

**Reshape Matrix**

```{r}
M <- matrix(c(3:14), nrow = 4)
print(M)
print("Matrix dimensions are ")
print(dim(M))
dim(M) = c(2,6)
print("New dimensions are")
print(dim(M))
# NOT Allowed 
#dim(M) = c(2,2) # Less values in target
#print("New dimensions are")
#print(dim(M))
#dim(M) = c(1,14) # More values in target
#print("New dimensions are")
#print(dim(M))
M
```

**Transpose of Matrix**

```{r}
M <- matrix(c(3:14), nrow = 4)
print(M)

print("transpose is")
t(M)
```


**Removing Elements from Matrix**

```{r}
M <- matrix(c(3:14), nrow = 4)
print(M)

print("Removing  third row")
M[-3,]
print("Removing  third & 4 th row")
M[c(-3,-4),]

# Original matrix is not modified!!
print("Removing  Second column")
M[,-2]
print("Removing  1 &3 column")
M[,c(-1,-3)]
print("Removing one row and one column ")
M[-1,-2]
```

**Seach An element in Matrix**
 Find index of an element in matrix
 
```{r}
M <- matrix(c(3:14), nrow = 4)
print(M)
print("Index of 9")
which(M == 9)
print("Elements excluding 9")
M[-which(M == 9)] # remove single element from matrix 
print("Index of even nos")
which(M %% 2 ==0)
print("Elements excluding even nos")
M[- which(M %%2 ==0)] # remove all even nos from the matrix
```


**Matrix Operations**

```{r}
M1 <- matrix(c(1:9), nrow = 3)
M2 <- matrix(c(11:19), nrow = 3)
cat("\n After addition \n")
print(M1+M2)
cat("\n After Substraction \n")
print(M1-M2)
cat("\n After elementwise Multiplication \n")
print(M1*M2)
cat("\n After Matrix Multiplication \n")
print(M1 %*% M2)
cat("\n After Division \n")
print(M1/M2)
cat("\n After Matrix Multiplication with scalar (constant)\n")
print(M1 * 4)

```

