---
title: "R Operators"
output: html_notebook
---

**Types of Operators**
1 Arithmetic Operators
2 Relational Operators
3 Logical Operators
4 Assignment Operators
5 Miscellaneous Operators

**Arithmetic Operators**
+ . - . * , /, %% (mod), %/% (integer division), ^

```{r}
x = c(15,22)
y = c(3,5)
x
y
cat("\n Addition :",x+y)
cat("\n Substraction :",x-y)
cat("\n Multiplication :", x*y)
cat("\n Division :", x/y)
cat("\n Mod (remainder):",x%%y)
cat("\n Integer division",x%/%y)
cat("\n Power operation:",x^y)
```

**Relational Operators**

< , > , == , !=, >=, <=

```{r}
x = c(15,22)
y = c(3,5)
cat("\n Less than ::",x<y)
cat("\n Greater than::",x>y)
cat("\n Equal ::", x==y)
cat("\n Not equal ::", x!=y)
cat("\n Greater than equal::",x>=y)
cat("\n Less than equal ::",x<=y)
```

**Logical Operators**

vector operator :: &&, || (Compares only first element of vectors)
Element wise operators :: &, | , ! (Compares all elements one by one)

Non Zero values are TRUE
[Negative nos are TRUE and positive nos are TRUE]

Zero is FALSE


```{r}
v <- c(3,0,56.23,2+2i)
z= c(TRUE,FALSE,TRUE,FALSE)
t <- c(0,3,TRUE,2+3i)
v
t
print(v&t)
print(v&&t)
#print(v||t)
#print(v|t)

x=10
y=0
print(x&y)
print(x&&y)

```

```{r}
# WAP to check working of & , &&, |, ||
x=c(23,0,34,-4,6)
y=c(23,67,12,78,5.67)
z=c(53,987,2.2,6,8)
#& , && , |, ||

```


**Assignment Operators**
<-, =, <<- :: left assignment
->, ->> :: right assignment

```{r}
x = 10
y<-20
z<<-24

34-> n
35 ->> b
cat(x,y,z,n,b)
```


**Miscellaneous Operators**

%in% :: This operator is used to identify if an element belongs to a vector. %in% returns a logical vector as long as the first argument, with a TRUE if that value can be found in the second argument and a FALSE otherwise
%*% :: This operator is used to get dot product of matrix.

```{r}
v1 <- 8
v2 <- 15
x=c(8,15)
t <- c(9,12,45,8,67,5,8)
print(v1 %in% t) 
print(v2 %in% t)
print(t %in% v1)
print(t %in% v2)
```

```{r}
x=c(8,15)
t <- c(9,12,15,8,67,5,8)
print(t %in% x)
print(which(t %in% x))
```

```{r}
x=c(8,15)
t <- c(9,12,15,8,67,5,8)
print(x %in% t)
print(which(x %in% t))
```


```{r}
x=c(1:10)
y=c(3,9,11,15,2,7,8,2)
x
y
x%in%y
x[x%in%y] #--> select all elements from x which are present in y 
y%in%x
y[y%in%x] # --> select all elements of y which are present in x
```


```{r}
M = matrix( c(2,6,5,1,10,4), nrow=2,ncol=3)
M1 = matrix( c(20,60,50,1,10,4), nrow=3,ncol=2)
print(M)
print(t(M))
ans = M * M
print(ans)
ans = M %*% t(M) #size 2 by 2
print(ans)
ans = M %*% M1 #size 2 by 2
print(ans)
```

```{r}

```

